Automated data-driven symbolic modeling of the lift forces acting on flapping wings

### 2021-02-01 paper/main.tex
Label columns as fitted and not fitted
Add TODO
Explain the difference between fitted and not fitted results

\subsection*{Data-driven models}
The strategy for modeling lift forces symbolically was to train candidate
equations on 58 of the 68 total experiments simultaneously, with the remaining
ten reserved for testing.  These 58 training sets were intended to represent a
diverse sampling of wing geometry, flapping speed and amplitude combinations.
The results from these 58 experiments were compiled into a single dataset
listing roughly two periods of smoothed, registered flapping data per
experiment.  The data consisted of instantaneous measured vertical force along
with the measured value of each kinematic variable, the first and second
derivatives of those variables, and the geometric parameters necessary to
describe the wing's shape and distribution of area, totaling 31 separate
variables.

By training on 58 experiments simultaneously, the symbolic regression algorithm
generated equations that fit all of the data, using the some, but not all, of
the provided variables to account for variations in kinematics and wing geometry
between experiments.   Table~\ref{table:eqmodels} lists five of the leading
candidate models, which were trained using mean absolute error (MAE) as the
fitness metric during the search. The search was conducted for approximately 24
hours, using 60 cores on a computing cluster.

% TODO: Make sure to define all variables used in these tables.
% TODO: Figure out why we're calling it "U-squared".
\begin{table}[ht]
\small
\centering
\begin{tabular}{|l|l|r|r|r|}
\hline
  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & {\bf Fitted MAE (N)} & {\bf MAE (N)}\\
\hline
\(EQ _5\) & \(C_1RY_{CM}\dot{\psi}^2\ + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 12 & 8.79\e{-4} & 1.19\e{-3}\\
\(EQ _4\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 14 & 8.66\e{-4} & 1.18\e{-3}\\
\(EQ _3\) & \(C_1RY_{CM}\dot{\psi}^2\cos(\psi-\psi^2) + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 16 & 8.49\e{-4} & 1.14\e{-3}\\
\(EQ _2\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\dot{\phi}\dot{\psi}\psi - C_4R\ddot{\theta}\) & 19 & 8.15\e{-4} & 1.14\e{-3}\\
\(EQ _1\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\dot{\phi}\psi - C_4R\dot{\phi}\dot{\psi}\psi - C_5R\ddot{\theta}\) & 24 & 7.73\e{-4} & 1.12\e{-3}\\
\hline
\(U^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
\(U^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
\(U^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.02\e{-3} & 1.23\e{-3}\\
PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.01\e{-3} & 1.20\e{-3}\\
\hline
\end{tabular}
\caption{\label{table:eqmodels}Comparison of data-driven models and analytical
  equations of lift.  The models are sorted by equation size, which is
  calculated as the sum of arithmetic and trigonometric operators required to
  compute a force prediction from kinematic and geometric data.  MAE reports
  average mean absolute error (in Newtons of vertical force) for the ten test
  experiments.}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/mae_nofit}
\caption{Pareto plot of all data-driven (``EQ Models'') and analytical models
  using mean values for model coefficients. All data-driven models were smaller
  and more accurate than analytical models.\label{fig:pareto_mae}}
\end{figure}

>>>>

% TODO: Reword this section to not focus on the number 58 right off the bat.
\subsection*{Data-driven models}
The strategy for modeling lift forces symbolically was to train candidate
equations on 58 of the 68 total experiments simultaneously, with the remaining
ten reserved for testing.  These 58 training sets were intended to represent a
diverse sampling of wing geometry, flapping speed and amplitude combinations.
The results from these 58 experiments were compiled into a single dataset
listing roughly two periods of smoothed, registered flapping data per
experiment.  The data consisted of instantaneous measured vertical force along
with the measured value of each kinematic variable, the first and second
derivatives of those variables, and the geometric parameters necessary to
describe the wing's shape and distribution of area, totaling 31 separate
variables. By training on 58 experiments simultaneously, the symbolic regression
algorithm generated equations that fit all of the data, using the some, but not
all, of the provided variables to account for variations in kinematics and wing
geometry between experiments. The search was conducted for approximately 24
hours, using 60 cores on a computing cluster.

Table~\ref{table:eqmodels} lists the performance of five of the leading
candidate models, along with five analytical models for comparison. We measure
performance as the mean absolute error (MAE) between predicted and measured
vertical force in the 10 test set experiments withheld from training. Our
primary results, provided in the column labeled ``MAE Not Fitted'', use
coefficient values for all models that were optimized on the training set.
However, since the accuracy of the analytical models may depend on coefficients
that vary with wing shape, flapping speed, or other factors, we also provide
secondary results in the column labeled ``MAE Fitted'', which use coefficient
values that were fitted to the test set experiments.

% TODO: Make sure to define all variables used in these tables.
% TODO: Figure out why we're calling it "U-squared".
% TODO: Decide on better terms than "fitted" and "not fitted".
\begin{table}[ht]
\small
\centering
\begin{tabular}{|l|l|r|r|r|}
\hline
  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\
    Fitted\end{tabular} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\ Not Fitted\end{tabular}\\
\hline
\(EQ _5\) & \(C_1RY_{CM}\dot{\psi}^2\ + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 12 & 8.79\e{-4} & 1.19\e{-3}\\
\(EQ _4\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 14 & 8.66\e{-4} & 1.18\e{-3}\\
\(EQ _3\) & \(C_1RY_{CM}\dot{\psi}^2\cos(\psi-\psi^2) + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 16 & 8.49\e{-4} & 1.14\e{-3}\\
\(EQ _2\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\dot{\phi}\dot{\psi}\psi - C_4R\ddot{\theta}\) & 19 & 8.15\e{-4} & 1.14\e{-3}\\
\(EQ _1\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\dot{\phi}\psi - C_4R\dot{\phi}\dot{\psi}\psi - C_5R\ddot{\theta}\) & 24 & 7.73\e{-4} & 1.12\e{-3}\\
\hline
\(U^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
\(U^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
\(U^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.02\e{-3} & 1.23\e{-3}\\
PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.01\e{-3} & 1.20\e{-3}\\
\hline
\end{tabular}
\caption{\label{table:eqmodels}Comparison of data-driven models and analytical
  equations of lift.  The models are sorted by equation size, which is
  calculated as the sum of arithmetic and trigonometric operators required to
  compute a force prediction from kinematic and geometric data.  MAE reports
  average mean absolute error (in Newtons of vertical force) for the ten test
  experiments.}
\end{table}

@@

% TODO: Remove all occurrences of the word "pareto" from the text, figures and filenames.
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/mae_nofit}
\caption{Pareto plot of all data-driven (``EQ Models'') and analytical models
  using mean values for model coefficients. All data-driven models were smaller
  and more accurate than analytical models.\label{fig:pareto_mae}}
\end{figure}
